Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals

### 2021-11-21 09_Article/01_Revision2_Reviewers/Response to Reviewers_00.tex

\title{Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals}
\author{Mathieu Simon, Michael Indermaur, Denis Schenk, Seyedmahdi Hosseinitabatabaei, Bettina	M. Willie and Philippe Zysset}
\journal{Bone}
\doi{BONE-D-21-00758-R1}

\begin{document}

\maketitle

\vspace{1em}Once again, the authors would like to thank the reviewers for helping us to improve our article. We do believe that their time spent working on this document will definitively improve its message to the scientific community. As for the first revision, answers to comments are provided point by point in this document. In the revised manuscript, the added elements are written in blue and the canceled text is in red.

@@

Page 9, second column, Discussion, lines 50-52
\begin{quote}
	The linear regression plot (Figure 5b) shows that when the stiffness term decreases to \Add{1 MPa (}10\textsuperscript{0}\Add{)} \Del{MPa} and lower, the linear regression tends to overestimate the stiffness.
\end{quote}

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\textwidth]
	{Pictures/02_GR_OI_LMM}
	\caption{Linear regression results for original OI group using the fixed effects of the linear mixed-effect model.}
	\label{CV_BVTV2}
\end{figure}

\newpage
\section{Clinical refocusing}
\RC Both reviewers were concerned about the fit of the manuscript content with the audience of Bone. While I agree that the Bone audience is accustomed to technical papers, please consider further refocusing to the clinical aspect. I would question where it is necessary to explicitly include equations 2 and 3 (page 4) since those are presumably outlined in the cited reference. If needed at all, considering moving to Appendix. Furthermore, equation 7 is not explained and therefore its relevance (and frequent reference later in the Discussion) is not clear.

\AR We understand your concern and agree to further refocusing to the clinical aspect. Regarding equations 2 and 3, even if they are outlined in cited references, we think that they are necessary for technical completeness. On the other hand, as they are more related to technical aspect we moved equation 2 and 3 but also equations 1,6 and 7 in the Appendix, as proposed. The text citing equation 7 was modified to describe the term contained and the signification of this NE, see below. We hope that these steps will help to decrease the technical difficulty and increase the clarity of the paper from a clinical point of view.

Page 5, second column, Methods 2.5.3, last paragraph
\begin{quote}
	Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE)\Del{, see Equation 6 and 7}. \Add{This relative norm error allows to quantify the accuracy of the prediction. Details of the computation of these numbers are available in Appendix A.}
\end{quote}

\section{Results validity}
\RC Most importantly, both reviewers had concerns about the validity of the results across the entire range of trabecular bone microarchitecture encountered in OI, and in particular the case where filtering was required to exclude highly heterogeneous bone microarchitecture. While it is understood that the focus of the paper is to explore whether fabric-elasticity relationships in OI and healthy bone are similar (thus justifying why filtering was performed), it is concluded that this is “encouraging the use of HR-pQCT based hFE outcomes for fracture risk assessment.” The comment about future fracture prediction by hFE is of course very important clinically, but the messaging about how to use these results are still not clear. Both the Abstract and the Discussion should comment on the implications of the fabric-elastic relationship not being defined for highly heterogeneous bone.

\AR We believe we can resolve these concerns. The filtering is made due to the sampling of cubic ROIs for computational homogenisation and does not filter bone microstructure from another. In our study, high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI and correlates with low, sometimes extremely low BV/TV and stiffness. That is why many more OI than healthy ROIs are filtered out.
The mentioned limitation resides in the sampling of homogenous cubes and the application of computational homogenisation, not in the fabric-elasticity relationship which application to heterogeneous ROIs is well-defined, smooth in BV/TV, smooth in fabric and independent of heterogeneity. Finally, the use of the fabric-elasticity relationship was validated experimentally also in osteoporotic bones where medullary cavities are also important.

\newpage
\RC Firstly, given there is only one OI type III where ROIs could be extracted, does this mean hFE should be restricted to OI type I? What evidence is there that it can be employed for OI type III or IV? This needs to be clearly justified.

\AR Although no further statistical analysis is meaningful due to the small numbers, ROIs from all OI types fit nicely on a unique fabric-elasticity relationship. The lack of homogeneous type III ROIs is simply due to the lack of bone in which case no relationship is needed. No outlier and no evidence of a different relationship was seen in our results for any OI type including the sample with OI type III.

\RC Second, in the future, when significant heterogeneity is encountered in a bone, do you still recommend to use hFE? In other words, please comment whether the findings of skewness for high heterogeneity affects if/when hFE is appropriate, and if there are alternatives to hFE in that case.

\AR The fabric-elasticity relationships between healthy and OI ROIs were found to be equivalent from a few MPa up to several GPa. Any ROI stiffness below a few MPa is meaningless in a mechanical analysis because of being several orders of magnitude smaller than the other ROIs and the cortex. Accordingly, hFE can be applied to bone with large voids or cavities as the BV/TV at the edges are smeared over the neighboring elements.

%Highly heterogeneous bone is rare in healthy conditions leading to the observed good correlations between simulations and experiments in the work of Varga et al. 2011, Hosseini et al. 2017, and Arias-Moreno et al. 2019. This fact is determinant. As low BV/TV will lead to low stiffness, the influence of such ROIs is negligible as it does not have a significant contribution to the stiffness of the full tibial section. %For example, the observed stiffness component $\lambda_{ii}$ of one highly heterogeneous ROI (worst case) is around 1 MPa and the predicted value is around 100 MPa. This apparently large error remains relatively small in comparison with homogeneous ROIs presenting a value approaching 10 GPa. 
%We believe that the heterogeneous ROIs are meaningless in the hFE analysis of a whole tibial section. In other words, a potential difference between healthy and OI for these

%Thus, only homogeneous ROIs (i.e. ROIs with higher BV/TV and connected to the surrounding bone) contribute significantly to the overall bone stiffness. %Moreover, despite the contribution of trabecular bone, the cortical bone will contribute even more significantly to tibial stiffness. Cortical bone having a higher BV/TV than trabecular bone, the relative error becomes even lower. 
%Another point is that ROIs here have 5.3mm side length. Since the element size is inferior to the ROI size, the heterogeneity is better resolved and the uncertainty affects a lower tissue volume. Finally, it is important to realise that the scatter associated with heterogeneous ROIs (as shown in Appendix) reflects a limitation of the homogenisation methodology to identify the fabric-elasticity relationship and is not a true error of the fabric-elasticity paradigm. The extrapolation of the fabric-elasticity relationship to heterogeneous ROIs that is made in hFE may be quite reasonable, but the actual uncertainty cannot be quantified with the available tools.

To improve the clarity of the message regarding validity of hFE among the OI types or heterogeneous bone, the manuscript was modified as follow:

Abstract, paragraph 4, last sentence
\begin{quote}
	The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology. \Add{However, this limitation does not challenge the fabric-elasticity relationship, which extrapolation to heterogeneous ROIs may be reasonable but can simply not be evaluated with the employed homogenisation methodology. Moreover, due to their low BV/TV, the potential (unknown) errors on these heterogeneous ROIs would have little influence on whole bone stiffness in comparison to homogeneous ROIs which are orders of magnitude stiffer.}
\end{quote}

Discussion, page 9, first column, line 56-59
\begin{quote}
	The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter. \Add{This comes from the fact that high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI.}
\end{quote}

Discussion, page 9, second column, line 55-60
\begin{quote}
	Some ROIs with low BV/TV do not have every side of the cube connected by bone, leading to extremely low terms in the stiffness tensor (see Appendix C). Trying to homogenise such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure 5b). \Add{This represents a limitation of the homogenization methodology and t}\Del{T}herefore, a filtering \Add{according to the heterogeneity level} is indispensable to assess and compare fabric-elasticity relationships, as done by Panyasantisuk et al. [28].
\end{quote}

\newpage
Discussion, page 10, second column, line 30
\begin{quote}
	\Add{Nevertheless, it should be noted that this relation between BV/TV and CV as well as the fabric-elasticity relationships are independent of the OI type. Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix D.}
\end{quote}

Discussion, page 10, second column, line 47-50
\begin{quote}
	These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships. \Add{Indeed, this confirms that filtering according to heterogeneity level allows to compare fabric-elasticity relationships of homogeneous ROIs where homogenization procedure is valid.}
\end{quote}

Discussion, page 10, second column, last paragraph, last sentence
\begin{quote}
	Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones\Add{, regardless the OI type}. Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness\Del{.}\Add{ order of magnitude lower than homogeneous ones and the cortex (< 1MPa compared to 10 GPa). Moreover, studies validated the use of fabric-elasticity relationships in osteoporotic bone [35,17] where medullary cavities are important and therefore lead to locally high heterogeneity.}
\end{quote}

Discussion, page 11, second column, line 40-24
\begin{quote}
	Moreover, having only one patient with OI type III where we could extract ROIs does not allow to perform statistics on OI types. \Add{Nevertheless, the lack of homogeneous type III ROIs is simply due to the lack of bone in which case no relationship is needed. No outlier and no evidence of a different relationship was seen in our results for any OI type including the sample with OI type III.}
\end{quote}

### 2021-11-21 09_Article/01_Revision2_Reviewers/Simon_2021_Bone_Manuscript.tex
Implement Philippe suggestions into manuscript

		Compared to healthy controls, we found the OI samples to have significantly lower BV/TV and trabecular number (Tb.N.), significantly higher CV, trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.). These results are in agreement with previous studies. The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology. \Add{Nevertheless, the error performed on these highly heterogeneous ROIs is negligible in comparison with homogeneous ROIs which are order of magnitude stiffer.}

		The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and led to similar correlation coefficients for both OI and healthy groups. Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals.\Add{ Moreover, the filtering step did not exclude a particular OI type.} Compared to previous studies, the stiffness constants from the 61 \si{\micro}m resolution HR-pQCT ROIs were lower than for the 36 \si{\micro}m resolution \si{\micro}CT ROIs.

		In conclusion, OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and \Del{less dense} \Add{have a lower BV/TV} than healthy controls. Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis.  Accordingly, the elastic properties used for FEA of healthy bones are also valid for OI bones.
	\end{abstract}

@@

	\subsection{HR-pQCT}
	HR-pQCT imaging (XtremeCTII, SCANCO Medical, Br\"{u}tisellen,
	Switzerland) was performed at the distal tibia on all patients from both groups. The distal tibia was chosen because numerous studies were performed on the distal radius \cite{AriasMoreno2019,Boutroy2008,Gross2013,Hosseini2017,Kocijan2015,Varga2011}, some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) \cite{Boutroy2005,Folkestad2012} but only one study was performed on the distal tibia with XtremeCTII \cite{Schenk2020}. Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}. Name\-ly, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}. On the other hand, for member of the OI group, the reference line is placed at the subchondral endplate of the ankle joint.  The mid-point of the scanned volume of interest was located at a distance of 7\% of total tibia length from the reference line \cite{GHASEMZADEH2017,Bonaretti2017}, see Figure \ref{01_OI}. Then, the scans were acquired according to the manufacturer's standard in vivo acquisition protocol. The scanned regions according to these two different protocols coincide approximately and are shown side by side in Figure \ref{01_ClinicalSections}. For both the healthy and the OI groups, each stack consisted of 168 slices and an isotropic voxel resolution of 61 \si{\micro}m. This led to a thickness of roughly 10.2 mm for each stack. Standardized scanning settings were used (voltage of 60 kVp, 900 $\mu$A, 100 ms integration time) for the healthy group as well as for the OI group. For the healthy group, motion artefacts of first, middle and last slice (i.e slices number 1, 252, and 504) were graded on a scale proposed by \citeauthor{Pialat2012} \cite{Pialat2012} of 1 (no motion artefacts) to 5 (extreme motion artefacts). The final grade of each scan was defined as the highest slice grade. For the OI group, as the scan consisted of one stack, only one grade was attributed using the same scale as for the control group. Scans were then processed independently of their quality grading. A summary of the scans grading is shown in Figure \ref{01_MotionArtefacts}.

@@

	In general, six cubic ROIs were selected at random position in each scan in a predefined area. Each ROI had to contain trabecular bone, but no cortical bone. For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number 1) to be at roughly the same anatomical location as for the OI group. Then, for both the healthy group and the OI group, the stack was divided into two halves with a transverse plane and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}. For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no\Add{t} enough trabecular bone. This led to 720 healthy ROIs from 120 individuals and 294 OI ROIs from 49 individuals.\\

@@

	After ROI cleaning, i.e. deletion of bone material regions unconnected to the main structure, the morphological analysis of ROIs was performed using medtool (v4.5; Dr. Pahr Ingenieurs e.U., Pfaffst\"{a}tten, Austria). The morphological parameters analyzed were: BV/TV, structural model index (SMI), trabecular number (Tb.N.), trabecular thickness (Tb.Th.), trabecular separation (Tb.Sp.), and the standard deviation of the trabecular separation (Tb.Sp.SD). Moreover, ROI tissue mineral density (TMD) and fabric was evaluated. The fabric tensor $\mathbf{M}$ was computed using mean intercept length (MIL) method \cite{Moreno2014}. \Del{It is a positive-definite second-order tensor computed as shown below.}\\

	\Del{where $m_i$ are the eigenvalues of $\mathbf{M}$ and $\mathbf{M}_i$ are the dyadic product of the corresponding eigenvectors $\mathbf{m}_i$ }\cite{Cowin1985,Harrigan1985}\Del{. The fabric tensor is independent of BV/TV and normalised with $tr(\mathbf{M}) = 3$.} The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI by dividing the highest eigenvalue by the lowest one. Figure \ref{01_FabricExample} shows an example of a typical ROI with the visualisation of its fabric tensor.\\

@@

	A \si{\micro}FE analysis was performed using \textsc{ABAQUS 6.14}. In brief, each voxel of the cleaned ROI was converted to a fully integrated linear brick elements (C3D8) using a direct voxel conversion approach. Then, a Young's modulus $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned to each element. The homogenisation process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}. Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require to rotate the ROI into the respective fabric coordinate system. Such rotation would potentially decrease image quality due to interpolation. The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}. This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV \Del{or $\rho$}, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$\Del{, as shown in Equation \ref{Eq202}}. \Add{This stiffness tensor is then transformed into the fabric coordinate system} and projected onto orthotropy, leading to 12 components. \Add{The technical details of the computations performed are presented in Appendix \ref{A0}.}\\

@@

	\subsection{Linear Regression}
	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model using the entire sample of both groups. Standard linear models assume independent and identically distributed (iid) variables. As this assumption was violated by the fact that six ROIs were analysed per individual, a linear mixed-effect model was preferred. This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, considered the non-independence of ROIs from the same individual. A more detailed form of this model is presented in Appendix \ref{A1}.

@@

	The linear regression was performed using the \textsc{statsmodels} package from \textsc{Python 3.6}. Linear regression quality for the TMD analysis was assessed using the Pearson correlation coefficient squared ($R^2$) and the standard error of the estimate (SE). Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE)\Del{, see Equation \ref{Eq205} and \ref{Eq206}}. \Add{This relative norm error allows to quantify the accuracy of the prediction. Details of the computation of these numbers are available in Appendix \ref{A0}.}

@@

	As the previous studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} have the same age range as our matched groups, we can compare the morphological parameters. The CT imaging system explains most of the differences between the absolute morphological values of the present study compared to the others. \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have performed their measurements on first generation XCT scanners with a voxel size of 82 \si{\micro}m, while we have used a second generation XCT with a voxel size of 61 \si{\micro}m. The work from \citeauthor{Agarwal2016}\cite{Agarwal2016} investigated differences between the two scanner types. They showed that BV/TV, Tb.Th., and Tb.Sp. are higher in second generation XCT scanners and Tb.N. is lower compared to first generation XCT. These findings are in full agreement with our observations. Another potential bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length. This influences the Tb.N. and Tb.Sp. as they depend on the ROI size. Moreover, conditions imposed for random ROI selection can lead to further biased values, especially for OI patients, as the ROI must contain at least some portion of trabecular bone. Even with the low sample size (2x28 individuals), the statistical tests have shown significant differences between groups, with the CV presenting the most significant difference. The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter. Finally, the significant differences observed in BV/TV and DA, even with matched age and sex, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values. \\

	The linear regressions performed in this study on original data sets showed $R^2_{adj}$ and NE in the expected range (i. e. slightly lower than \citeauthor{Gross2013}\cite{Gross2013} and \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}) for the healthy group. Components of the stiffness tensors are distributed to both sides of the diagonal. On the other hand, the linear regressions performed in this study using the original OI data set presents lower $R^2_{adj}$ and higher NE than comparable linear regressions in literature \cite{Gross2013,Panyasantisuk2015}. The important value of NE and its standard deviation shows that the fitted stiffness can deviate significantly from the observation. These differences come from ROIs presenting a low stiffness. The linear regression plot (Figure \ref{02_OI}) shows that when the stiffness term decreases to \Add{1 MPa (}10\textsuperscript{0}\Add{)} \Del{MPa} and lower, the linear regression tends to overestimate the stiffness. This is because ROI stiffness is highly dependent on BV/TV values. Some ROIs with low BV/TV do not have every side of the cube connected by bone, leading to extremely low terms in the stiffness tensor (see Appendix \ref{A2}). Trying to homogenise such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure \ref{02_OI}). Therefore, a filtering is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}. An alternative to CV filtering for assessing the ROI heterogeneity could be to compute the proportion of the area filled by bone on each of the six faces of the ROI. \\

	Figure \ref{02_CV_BVTV} presents the CV in relation to BV/TV. It shows a tendency of CV values to increase with decreasing BV/TV values. Effectively, if the quantity of material inside the ROI decreases, the distribution homogeneity of this mass is more sensitive and therefore can quickly become highly heterogeneous. A simple assumption about this relation is that it could be monotonic. Pearson's correlation coefficient being strictly negative confirms a negative monotonic relation. As some ROIs with higher BV/TV still present high CV values, imposing a fixed threshold for subsequent homogenisation seems feasible. However, the value of this threshold results from an optimisation process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be investigated further. \Add{Nevertheless, it should be noted that this relation between BV/TV and CV as well as the homogenization process (i.e. fabric-elasticity relationships) are independent of the OI type. Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix} \ref{A3}\Add{.} \\

>>>>

		Compared to healthy controls, we found the OI samples to have significantly lower BV/TV and trabecular number (Tb.N.), significantly higher CV, trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.). These results are in agreement with previous studies. The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology. However, this limitation does not challenge the fabric-elasticity relationship, which extrapolation to heterogeneous ROIs may be reasonable but can simply not be evaluated with the employed homogenisation methodology. Moreover, due to their low BV/TV, the potential (unknown) errors on these heterogeneous ROIs would have little influence on whole bone stiffness in comparison to homogeneous ROIs which are orders of magnitude stiffer.

		The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and led to similar correlation coefficients for both OI and healthy groups. Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals. Moreover, the filtering step did not exclude a particular OI type. Compared to previous studies, the stiffness constants from the 61 \si{\micro}m resolution HR-pQCT ROIs were lower than for the 36 \si{\micro}m resolution \si{\micro}CT ROIs.

		In conclusion, OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and have a lower BV/TV than healthy controls. Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis.  Accordingly, the elastic properties used for FEA of healthy bones are also valid for OI bones.
	\end{abstract}

@@

	\subsection{HR-pQCT}
	HR-pQCT imaging (XtremeCTII, SCANCO Medical, Br\"{u}tisellen,
	Switzerland) was performed at the distal tibia on all patients from both groups. The distal tibia was chosen because numerous studies were performed on the distal radius \cite{AriasMoreno2019,Boutroy2008,Gross2013,Hosseini2017,Kocijan2015,Varga2011}, some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) \cite{Boutroy2005,Folkestad2012} but only one study was performed on the distal tibia with XtremeCTII \cite{Schenk2020}. Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}. Na\-me\-ly, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}. On the other hand, for member of the OI group, the reference line is placed at the subchondral endplate of the ankle joint.  The mid-point of the scanned volume of interest was located at a distance of 7\% of total tibia length from the reference line \cite{GHASEMZADEH2017,Bonaretti2017}, see Figure \ref{01_OI}. Then, the scans were acquired according to the manufacturer's standard in vivo acquisition protocol. The scanned regions according to these two different protocols coincide approximately and are shown side by side in Figure \ref{01_ClinicalSections}. For both the healthy and the OI groups, each stack consisted of 168 slices and an isotropic voxel resolution of 61 \si{\micro}m. This led to a thickness of roughly 10.2 mm for each stack. Standardized scanning settings were used (voltage of 60 kVp, 900 $\mu$A, 100 ms integration time) for the healthy group as well as for the OI group. For the healthy group, motion artefacts of first, middle and last slice (i.e slices number 1, 252, and 504) were graded on a scale proposed by \citeauthor{Pialat2012} \cite{Pialat2012} of 1 (no motion artefacts) to 5 (extreme motion artefacts). The final grade of each scan was defined as the highest slice grade. For the OI group, as the scan consisted of one stack, only one grade was attributed using the same scale as for the control group. Scans were then processed independently of their quality grading. A summary of the scans grading is shown in Figure \ref{01_MotionArtefacts}.

@@

	In general, six cubic ROIs were selected at random position in each scan in a predefined area. Each ROI had to contain trabecular bone, but no cortical bone. For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number 1) to be at roughly the same anatomical location as for the OI group. Then, for both the healthy group and the OI group, the stack was divided into two halves with a transverse plane and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}. For one individual diagnosed with OI type III it was not possible to extract any ROI as there was not enough trabecular bone. This led to 720 healthy ROIs from 120 individuals and 294 OI ROIs from 49 individuals.\\

@@

	After ROI cleaning, i.e. deletion of bone material regions unconnected to the main structure, the morphological analysis of ROIs was performed using medtool (v4.5; Dr. Pahr Ingenieurs e.U., Pfaffst\"{a}tten, Austria). The morphological parameters analyzed were: BV/TV, structural model index (SMI), trabecular number (Tb.N.), trabecular thickness (Tb.Th.), trabecular separation (Tb.Sp.), and the standard deviation of the trabecular separation (Tb.Sp.SD). Moreover, ROI tissue mineral density (TMD) and fabric was evaluated. The fabric tensor $\mathbf{M}$ was computed using mean intercept length (MIL) method \cite{Moreno2014}. The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI by dividing the highest eigenvalue by the lowest one. Figure \ref{01_FabricExample} shows an example of a typical ROI with the visualisation of its fabric tensor.\\

@@

	A \si{\micro}FE analysis was performed using \textsc{ABAQUS 6.14}. In brief, each voxel of the cleaned ROI was converted to a fully integrated linear brick elements (C3D8) using a direct voxel conversion approach. Then, a Young's modulus $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned to each element. The homogenisation process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}. Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require to rotate the ROI into the respective fabric coordinate system. Such rotation would potentially decrease image quality due to interpolation. The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}. This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$. This stiffness tensor is then transformed into the fabric coordinate system and projected onto orthotropy, leading to 12 components. The technical details of the computations performed are presented in Appendix \ref{A0}.\\

@@

	\subsection{Linear Regression}
	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model using the entire sample of both groups. Standard linear models assume independent and identically distributed (iid) variables. As this assumption was violated by the fact that six ROIs were analysed per individual, a linear mixed-effect model was preferred. This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, considered the non-in\-de\-pen\-dence of ROIs from the same individual. A more detailed form of this model is presented in Appendix \ref{A1}.

@@

	The linear regression was performed using the \textsc{statsmodels} package from \textsc{Python 3.6}. Linear regression quality for the TMD analysis was assessed using the Pearson correlation coefficient squared ($R^2$) and the standard error of the estimate (SE). Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE). This relative norm error allows to quantify the accuracy of the prediction. Details of the computation of these numbers are available in Appendix \ref{A0}.

@@

	As the previous studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} have the same age range as our matched groups, we can compare the morphological parameters. The CT imaging system explains most of the differences between the absolute morphological values of the present study compared to the others. \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have performed their measurements on first generation XCT scanners with a voxel size of 82 \si{\micro}m, while we have used a second generation XCT with a voxel size of 61 \si{\micro}m. The work from \citeauthor{Agarwal2016}\cite{Agarwal2016} investigated differences between the two scanner types. They showed that BV/TV, Tb.Th., and Tb.Sp. are higher in second generation XCT scanners and Tb.N. is lower compared to first generation XCT. These findings are in full agreement with our observations. Another potential bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length. This influences the Tb.N. and Tb.Sp. as they depend on the ROI size. Moreover, conditions imposed for random ROI selection can lead to further biased values, especially for OI patients, as the ROI must contain at least some portion of trabecular bone. Even with the low sample size (2x28 individuals), the statistical tests have shown significant differences between groups, with the CV presenting the most significant difference. The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter. This comes from the fact that high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI. Finally, the significant differences observed in BV/TV and DA, even with matched age and sex, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values. \\

	The linear regressions performed in this study on original data sets showed $R^2_{adj}$ and NE in the expected range (i. e. slightly lower than \citeauthor{Gross2013}\cite{Gross2013} and \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}) for the healthy group. Components of the stiffness tensors are distributed to both sides of the diagonal. On the other hand, the linear regressions performed in this study using the original OI data set presents lower $R^2_{adj}$ and higher NE than comparable linear regressions in literature \cite{Gross2013,Panyasantisuk2015}. The important value of NE and its standard deviation shows that the fitted stiffness can deviate significantly from the observation. These differences come from ROIs presenting a low stiffness. The linear regression plot (Figure \ref{02_OI}) shows that when the stiffness term decreases to 1 MPa (10\textsuperscript{0}) and lower, the linear regression tends to overestimate the stiffness. This is because ROI stiffness is highly dependent on BV/TV values. Some ROIs with low BV/TV do not have every side of the cube connected by bone, leading to extremely low terms in the stiffness tensor (see Appendix \ref{A2}). Trying to homogenise such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure \ref{02_OI}). This represents a limitation of the homogenization methodology and therefore, a filtering according to the heterogeneity level is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}. An alternative to CV filtering for assessing the ROI heterogeneity could be to compute the proportion of the area filled by bone on each of the six faces of the ROI. \\

	Figure \ref{02_CV_BVTV} presents the CV in relation to BV/TV. It shows a tendency of CV values to increase with decreasing BV/TV values. Effectively, if the quantity of material inside the ROI decreases, the distribution homogeneity of this mass is more sensitive and therefore can quickly become highly heterogeneous. A simple assumption about this relation is that it could be monotonic. Pearson's correlation coefficient being strictly negative confirms a negative monotonic relation. As some ROIs with higher BV/TV still present high CV values, imposing a fixed threshold for subsequent homogenisation seems feasible. However, the value of this threshold results from an optimisation process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be investigated further. Nevertheless, it should be noted that this relation between BV/TV and CV as well as the homogenization process (i.e. fabric-elasticity relationships) are independent of the OI type. Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix \ref{A3}. \\

	The linear regressions performed on the filtered data sets present a direct effect of filtering as the ROIs meeting the homogeneity assumption lead to better results compared to linear regression including ROIs with high CV values. For the healthy group (N=603), the relatively small decrease of $R^2_{adj}$ (0.5\%) compared to the linear regression using the unfiltered data set is negligible. On the other hand, NE values are decreased by 2\% and therefore improved. The filtering eliminates data points further away from the diagonal (better NE) and other data points close to the diagonal leading to a smaller number of points (modifying $R^2_{adj}$). For the OI group, filtering leads to an important improvement of the linear regression (i.e. higher $R^2_{adj}$ and lower NE). $R^2_{adj}$, NE, and the range of stiffness values are almost at the level of the healthy group. These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships. Indeed, this confirms that filtering according to heterogeneity level allows to compare fabric-elasticity relationships of homogeneous ROIs where homogenization procedure is valid.\\
